---
title: "Getting Started with mvGWAS"
author: "Robert W. Corty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Now that you've installed and loaded the `mvGWAS` package [(instructions here)](https://github.com/rcorty/mvGWAS), let's run through a typical workflow.

## Create an mvGWAS object

To create an `mvGWAS` object, you'll need two things on your filesystem:

1.  A directory with some .VCF or .VCF.GZ files in it.
For now, set it up so there are just a few (<10) files each of which has only a few variants (<1000) on a few participants (<100).
2.  A file with some phenotype data, either a .RDS file or a .CSV file.  
The first column of the phenotype data _must_ be called 'ID' and contain an ID for each individual.
This is the ID that will be used to match phenotype to the genotypes in the VCF files, so be sure you're using the same IDs.


```{r example, eval=FALSE}
library(mvGWAS)

my_first_gwas <- mvGWAS$new(phenotype_file = 'my_phenotype_file.csv',
                            genotype_directory = 'genotype_dir')
```

The first thing to note here is that we didn't define `mvGWAS`.
That's because it's defined in the package `mvGWAS` and exported.
It is a "reference object generator".
When it's used as above, it generates a reference object.

We've saved the reference object it generated as `my_first_gwas`.
This object contains all the methods for conducting an mvGWAS.
So, going forward, `my_first_gwas` will be where all the action is.



## Conduct a 'genome' scan on the local machine

We'll imagine that the handful of VCF files in `genotype_dir` represent the genome.
Since there's not much data, we can conduct the whole scan on the local machine.
We'll use the `conduct_scan` function, which takes two arguments: `mean_formula` and `var_formula`.
They use the [model formula system](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/formula.html) that is typical in R.

`mean_formula` is a two-sided formula.
Its right side is just the phenotype to be analyzed.
Its left side is covariates and keywords, separated by `+`.
There are three keywords available:



1.  DS -- the expected alternate allele dosage.
From a linear model perspective, this generates a one-column numeric design matrix where entry $i$ is the expected alternate allele dosage of individual $i$.
2.  GT -- the most likely genotype, a factor with up to 3 levels.
From a linear model perspective, this generates a two-column binary design matrix with homozygote reference as the reference level.
Entry [$i$,1] is 1 if the most likely genotype of individual $i$ is a heterozygote and 0 otherwise.
Entry [$i$,2] is 1 if the most likely genotype of individual $i$ is homozygous for the alternate allele and 0 otherwise.
3.  GP -- the probability of each genotype.
From a linear model perspective, this generates a two-column numeric design matrix with complete certainty of homozygote reference as the reference level.
Entry [$i$,1] is the probability that individual $i$ is a heterozygote.
Entry [$i$,2] is the probability that individual $i$ is homozygrous for the alternate allele.


`var_formula` is a one-sided formula.
It has no left side.
Its right side follows the same rules as `mean_formula`.


```{r eval=FALSE}
my_gwas$conduct_scan(mean_formula = my_trait ~ PC1 + PC2 + sex + age + DS,
                     var_formula = ~ PC1 + PC2 + sex + GT)
```

By default, `conduct_scan()` runs on the local machine, using as many cores as are available (based on `parallel::detectCores()`).
If you have a multi-core computer but don't want to use them all in `conduct_scan()`, you can pass in the maximum number of cores to use as `num_cores`.
For example:

```{r eval=FALSE}
my_gwas$conduct_scan(mean_formula = my_trait ~ PC1 + PC2 + sex + age + DS,
                     var_formula = ~ PC1 + PC2 + sex + GT,
                     num_cores = 2)
```

If `conduct_scan()` returns `TRUE`, it worked without error and the results are accessible as `my_gwas$results`.

## View the results

```{r eval=FALSE}
my_gwas$qq_plot()
my_gwas$manhattan_plot()
```

## View the joint null model

Because there are up to three hypotheses being tested, there are up to three null models.
Only the null model for the joint test


## Genomic control


## Write the results to file

```{r eval=FALSE}
aaabbb

```



